Literature Review

### 2017-06-13 Literature Review/A_Brief_Literature_Review.tex
Literature Review updated for final submission

Forecasting is a core element of the processes of a business. Producing
forecasts of high quality is essential as they play an important role in
inventory control, purchasing decisions of new equipment and investment
decisions, and are needed for effective planning (Granvik
\protect\hyperlink{ref-granvik2010forecasting}{2010}). Automatic
forecasting techniques have commonly been used in business to produce a
wide variety of forecasts and are a useful tool for producing a large
number of forecasts. Automatic techniques can be used by non-experts
without training in the use of time series (R. J. Hyndman, Khandakar,
and others \protect\hyperlink{ref-hyndman2007automatic}{2007}).
Automatic forecasting algorithms must have the ability to select the
appropriate model for forecasting and estimate model parameters, without
the need for human intervention (R. J. Hyndman, Khandakar, and others
\protect\hyperlink{ref-hyndman2007automatic}{2007}). The most common
automatic forecasting techniques used are the ARIMA and Exponential
Smoothing (ETS) models. ETS models have been used extensively in
forecasting demand for inventories (Billah et al.
\protect\hyperlink{ref-billah2006exponential}{2006}). They have
performed well in forecasting competitions against more complex models
and are particularly good at forecasting seasonal data over short
horizons (R. J. Hyndman et al.
\protect\hyperlink{ref-hyndman2002state}{2002}).

Automatic forecasting techniques can be too inflexible and do not allow
for analysts with specialized knowledge to incorporate their expertise
into the model effectively. This leads to difficulties in producing
forecasts of high quality, thus the demand for high quality forecasts
far exceeds the rate at which they are produced (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). In December 2016,
Facebook released their forecasting model, Prophet. According to Taylor
and Letham (\protect\hyperlink{ref-taylor2017forecasting}{2017}),
Prophet is a flexible model that can be adjusted by a large number of
non-experts who have little knowledge about time series, however have
expert knowledge on the data-generating process. Prophet allows for a
large number of forecasts to be produced across a variety of problems
and consists of a robust evaluation system that allows for a large
number of forecasts be evaluated and compared. This is Facebook's
definition of forecasting at scale, and provides a solution to the
problems that automatic forecasting pose (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\section{\texorpdfstring{Exponential State Space Models
\label{Exponential State Space Models}}{Exponential State Space Models }}\label{exponential-state-space-models}

\subsection{Exponential Smoothing
Methods}\label{exponential-smoothing-methods}

ETS methods originated in the 1950's and is the key work of Brown ,
Holt, and Winters (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). It has been widely used for
many applications in industry and has been the foundation of many
successful forecasting methods (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). Exponential smoothing
techniques produce forecasts by calculating them as a weighted average
of past observations. Observations that are most recent are more heavily
weighted, with these weights decaying exponentially as the observations
get older.

Exponential smoothing methods were originally classified by Pegel
according to the components present in the time series (trend and
seasonality), and whether they were additive or multiplicative in nature
(De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}). Gardner
and Taylor further extended this classification to include a damped
additive and multiplicative trend respectively (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). A taxonomy of exponential
smoothing methods was later created by Hyndman et al. and extended by
Taylor (De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}).
It considered all possible combinations of the trend and seasonality
components thus producing 15 methods of exponential smoothing as seen in
table \ref{tab1}.

\begin{table}[H]
\centering
\begin{tabular}{rlll}
  \hline
 & Trend &  & Seasonality \\ 
  \hline
 & None(N) & Additive(A) & Multiplicative(M) \\ 
  None(N) & (N,N) & (N,A) & (N,M) \\ 
  Additive(A) & (A,N) & (A,A) & (A,M) \\ 
  Additive\_damped(Ad) & (Ad,N) & (Ad,A) & (Ad,M) \\ 
  Multiplicative(M) & (M,N) & (M,A) & (M,M) \\ 
  Multiplicative\_damped(Md) & (Md,N) & (Md,A) & (Md,M) \\ 
   \hline
\end{tabular}
\caption{Taxonomy of Exponential Smoothing Methods \label{tab1}} 
\end{table}

\hfill

The most common of these methods are Simple Exponential Smoothing (N,N),
Holt's linear method (A,N) and Holt-Winters seasonal method (A,A).

Although exponential smoothing methods performed well and have been
widely used in business, they did not have any statistical framework
underpinning them (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). They produced point estimates
rather than prediction intervals and selection of a forecasting method
was specific to the time series in question. Efforts were made by Box
and Jenkins to create a statistical framework that underlies exponential
smoothing methods and they found that forecasts produced by linear
methods are special cases of ARIMA models (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). This however does not hold for
the multiplicative methods.

\subsection{State Space Models}\label{state-space-models}

The motivation behind state space models was the need for a statistical
framework that underlies all methods of exponential smoothing (De
Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}). R. J.
Hyndman et al. (\protect\hyperlink{ref-hyndman2002state}{2002}) provided
this framework as an extension of Ord, Koehler, and Snyder's class of
exponential state space models which underpinned some of the smoothing
methods (De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}).
He proposed two state space models, one corresponding to an additive
error and the other to a multiplicative error. By specifying a
distribution for these error terms, he turned the previously
deterministic methods into stochastic models. These two models were then
applied to all 15 smoothing methods, producing 30 models in total.

The two state space models produce identical point estimates to each
other as well as to the exponential smoothing methods, however they will
generate differing prediction intervals (R. Hyndman and Athanasopoulos
\protect\hyperlink{ref-hyndman2013forecasting}{2013}). Each model
consists of a measurement equation and a transition equation. The
measurement equation shows the relationship between the observations and
the unobserved components (level, trend seasonality), while the
transition equation shows how the unobserved components change over
time.

The introduction of the exponential state space model has allowed for
model selection using information criteria rather than using an ad hoc
method. It has also made the calculation of the likelihood easier and
simulation from the underlying model possible (R. J. Hyndman et al.
\protect\hyperlink{ref-hyndman2002state}{2002}).

\subsection{Application and
Performance}\label{application-and-performance}

In a paper by Smith and Agrawal
(\protect\hyperlink{ref-smith2015comparison}{2015}), Holt-Winters
exponential smoothing (HWES) model was applied to time series data
relating to data on patents. The patent data was classified into three
different groups; monthly forecasts were made and the results were
compared to forecasts produced by ARIMA models. To avoid
over-forecasting, the trend of the HWES was damped to produce forecasts
with a flat trend for later observations. The results showed that all
models forecasted the data adequately. The HWES model fitted the data
well and performed better than the ARIMA models in the case where the
time series for the patent group was fairly stationary. Based on the
forecasting results, it could not be concluded that a specific time
series model is better for forecasting patent data.

Hassani et al. (\protect\hyperlink{ref-hassani2015forecasting}{2015})
forecasted tourism demand in specific European countries over a short,
medium and long term horizon. The ETS model was used to produce
forecasts and was compared to six parametric and non-parametric
techniques. The results showed that no single model is best for
producing forecasts across all countries and time horizons. The ETS
model together with the Neural Networks (NN) and Fractionalized ARIMA
(ARFIMA) produced forecasts with the least amount of accuracy and are
thus not relevant models for forecasting tourism demand in Europe. 24
steps-ahead error forecasts produced by the ETS model were found to be
larger than the errors produced by the benchmark model - Recurrent
Single Spectrum Analysis (SSA-R) across all European countries.

\subsection{Robustness}\label{robustness}

Gardner (\protect\hyperlink{ref-gardner2006exponential}{2006}) explained
the robustness of ETS models by comparing them to equivalent models.
Simple exponential smoothing which is equivalent to an ARIMA(0,1,1)
model is the most robust method and has performed well in forecasting
many types of time series. By comparing simple exponential smoothing to
lower order ARIMA models, it is seen that errors that arise due to model
specification is not as much of a problem as ARIMA models. Furthermore,
ARIMA models tend to produce larger MSE due to model selection errors,
and this is worsened if the errors are not normally distributed (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

Satchell and Timmermann found that for time series with a finite
history, the weights assigned to each observation in the simple
exponential smoothing method are robust, provided that the variance of
the random walk relative to the variance of the error component is not
too small (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

Forecasts produced by simple exponential smoothing have performed well
in modelling annual sales of different products, as well as aggregated
economic series with low sampling frequencies (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

\section{Prophet}\label{prophet}

\subsection{Bayesian Generalized Additive
Model}\label{bayesian-generalized-additive-model}

Prophet is the key work of Taylor and Letham
(\protect\hyperlink{ref-taylor2017forecasting}{2017}) and consists of a
decomposable model with a component for growth, seasonality, and
holidays. These components consist of linear and non-linear functions of
time. This is similar to a Generalized Additive Model (GAM) which is a
regression model that consists of non-linear and linear regression
functions applied to predictor variables (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Prophet, like GAM,
frames the forecasting problem as a curve fitting exercise using
backfitting to find the regression functions.

The growth component is modelled in a similar way to modelling
population growths which uses a logistic growth model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Populations
typically grow non-linearly (although the growth component could also be
linear) up to an upper bound known as the carrying capacity, and remains
constant thereafter. The rate at which the population grows changes over
time, and this is accounted for by including changepoints in the model
where the growth rate may be changed (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). These changepoints
can be automatically selected and have a Laplace prior distribution
placed on them (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). The parameter of
the prior distribution can be used to adjust the growth rate of the
changepoints. This allows non-experts with knowledge about events that
may affect growth to use the parameter as a knob and adjust it to
increase or decrease the number of changepoints (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). It also allows for
the analyst to add changepoints which the automatic selection procedure
may have missed (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Furthermore,
analysts may also specify the carrying capacity and adjust it based on
their knowledge of the total market size (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

The decomposable form of the model allows for components to be easily
added to it. This allows for multiple seasonality components with
different periods to be added to the model. The seasonality component is
modelled by a Fourier series, with the parameters of the Fourier series
having a Normal prior distribution (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). The variance of
the parameters can be adjusted by analysts to smooth the model and
change how much of historical seasonality is projected to the future
(Taylor and Letham \protect\hyperlink{ref-taylor2017forecasting}{2017}).

The name, date, and country of past and future holidays and events may
be inputted by the analyst into a list (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). By specifying the
country in which the events take place or the holidays occur, separate
lists can be populated for global events/holidays and country-specific
events/holidays. The union of the two lists are then used for
forecasting. Like seasonality, a Normal prior distribution is placed on
the parameters of the holidays, and the variance of the parameters can
be adjusted by analysts to smooth the model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\subsection{Application and
Performance}\label{application-and-performance-1}

Taylor and Letham (\protect\hyperlink{ref-taylor2017forecasting}{2017})
forecasted the number of events on Facebook using Prophet. The time
series was impacted by holidays, had strong multi-period seasonality,
and a piecewise trend. The forecasts produced by Prophet were compared
to forecasts produced by common automatic forecasting techniques such as
ETS, ARIMA, and the seasonal naïve model, as well as to simple models
such as the naïve model. While the ETS and seasonal naïve model were
quite robust, the ARIMA forecasts were fragile. No model besides Prophet
accounted for the dips around holidays and the upward trend of the time
series towards later observations. Hence, Prophet had lower forecasting
errors compared to the automatic forecasting methods.

\subsection{Robustness}\label{robustness-1}

Prophet is a flexible model and has intuitive parameters that can be
easily interpreted by human beings. The GAM is fitted quickly, allowing
the analyst to interactively change the model parameters (Taylor and
Letham \protect\hyperlink{ref-taylor2017forecasting}{2017}). Unlike
ARIMA models, Prophet can produce forecasts over different scales and
allows for missing values in the time series without the need for
interpolation (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

When a large number of forecasts are produced, manually identifying
problematic forecasts becomes a time consuming and difficult task.
Prophet provides a semi-automated forecast evaluation system that
selects the best model that fits the data. When there are large forecast
errors, the forecasts are flagged so that the analyst can explore the
cause of the errors, identify and remove potential outliers and either
adjust the model or choose a more appropriate model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\section{Evaluation of Forecasts}\label{evaluation-of-forecasts}

\subsection{Common Methods}\label{common-methods}

Model selection has historically been very subjective, with no
techniques that offered mathematical rigour for choosing models (Smith
and Agrawal \protect\hyperlink{ref-smith2015comparison}{2015}). Progress
has been made over time and many mathematical techniques have been
introduced for model selection. Use of information criteria is one such
method, with Akaike's information criterion (AIC) and Bayes information
criterion (BIC) being the most common (Smith and Agrawal
\protect\hyperlink{ref-smith2015comparison}{2015}). These information
criteria minimize forecasting errors while penalizing it for
overfitting. Information criteria should be used to compare models of
similar structure only. To compare models of different structure, the
most commonly used measures of forecast errors are the Root Mean Squared
error (RMSE), Mean Absolute Error (MAE), Mean Absolute Percentage Error
(MAPE), and Mean Absolute Scaled Error (MASE). A lower value of these
statistics indicates that the forecasts are better, with a MASE less
than one indicating that the forecasts are better than those produced by
a naïve model (Smith and Agrawal
\protect\hyperlink{ref-smith2015comparison}{2015}).

Smith and Agrawal (\protect\hyperlink{ref-smith2015comparison}{2015})
have used these methods in their paper when evaluating patent forecasts
and making comparisons between the HWES and ARIMA models. In a paper by
Hassani et al. (\protect\hyperlink{ref-hassani2015forecasting}{2015})
they chose their ETS model based on the selection that R automatically
made for the time series. They compared models of different structure
using the RMSE and a modified Diebold-Mariano test. R. J. Hyndman et al.
(\protect\hyperlink{ref-hyndman2002state}{2002}) suggests that ETS
models should be selected using the AIC rather than the MSE or MAPE. The
AIC provides a way of choosing between models with an additive error and
models with a multiplicative error since it is based on the likelihood
function. The MSE and MAPE are not able to select between the two error
types because both models produce the same point estimates. According to
De Gooijer and Hyndman (\protect\hyperlink{ref-de200625}{2006}), the
best approach of selecting models with different structure is the
Diebold-Mariano test. The MSE is scale dependent and should not be used
to make comparisons while the MAPE encounters difficulties when the time
series values are close to or equal to zero.

\subsection{Mincer-Zarnowitz Approach}\label{mincer-zarnowitz-approach}

Mincer and Zarnowitz
(\protect\hyperlink{ref-mincer1969evaluation}{1969}) proposed two
methods of measuring forecast accuracy - an absolute and a relative
measure. Absolute measures of accuracy measure the distance between
actual and predicted values. If the mean distance is equal to zero, the
forecast is said to be unbiased (Mincer and Zarnowitz
\protect\hyperlink{ref-mincer1969evaluation}{1969}). To analyse absolute
forecast errors, the observed values are regressed against the predicted
values. A joint hypothesis test is performed to check if the intercept
is equal to zero and the slope is equal to one. If the intercept is
equal to zero, the forecasts are unbiased and do not overestimate or
underestimate the data. If the slope is equal to one, the residual
errors are uncorrelated with the predictions and the forecast is said to
be efficient (Mincer and Zarnowitz
\protect\hyperlink{ref-mincer1969evaluation}{1969}). The MSE is a
measure that can be decomposed into the residual variance and two
components, one reflecting the bias and the other reflecting the
inefficiency of the forecast. If the forecasts are unbiased and
efficient, the MSE will reduce to the residual variance (Mincer and
Zarnowitz \protect\hyperlink{ref-mincer1969evaluation}{1969}). Absolute
forecast measures cannot be used to make comparisons between forecasts
with different scales or economic variables. Thiel states that the
consequences of forecasting errors and how they impact decisions is of
greater importance than the size of the forecasting errors (Mincer and
Zarnowitz \protect\hyperlink{ref-mincer1969evaluation}{1969}).

Rather than using absolute measures of forecast accuracy, Mincer and
Zarnowitz (\protect\hyperlink{ref-mincer1969evaluation}{1969}) suggested
the use of relative measures of forecast accuracy. Relative accuracy
analysis allows for meaningful comparisons of different forecasting
methods to be made. They proposed an index which considers the ratio of
the MSE of the forecast to the MSE of a benchmark forecasting method.
They used an extrapolation of the data history as a benchmark, as it is
a cost effective and accessible method, however the benchmark may be any
method which is relevant for comparison. This ratio is known as the
Relative Mean Square Error (RM).

>>>>

The ability to forecast exchange rates is of great importance to policy
makers, businesses and speculators. Policy makers are interested in
forecasting exchange rates to assess how its movement will impact
inflation, and how monetary policy will need to be adjusted in the
future to keep inflation within the inflation target (Fahimifard et al.
\protect\hyperlink{ref-fahimifard2009}{2009}). Businesses need to assess
how future movements in exchange rates will affect the cost of importing
goods and the demand for goods exported, while speculators aim to
identify opportunities in price movements which can be exploited to make
a profit.

Producing accurate forecasts of exchange rates has proven to be a
difficult task due to the noise of the time series (Abu-Mostafa and
Atiya \protect\hyperlink{ref-abu1996}{1996}). Some economists argue that
the exchange rate market is very efficient and future movements in price
are unpredictable (Abu-Mostafa and Atiya
\protect\hyperlink{ref-abu1996}{1996}). This belief in the efficient
market hypothesis has led many researchers to model exchange rates using
a Random Walk (RW) model (Abu-Mostafa and Atiya
\protect\hyperlink{ref-abu1996}{1996}). Although the RW model has
produced superior forecasts, some researchers believe that future prices
depend on present and past events, and other techniques such as the
autoregressive integrated moving average (ARIMA) model has been applied
to forecast exchange rates (Abu-Mostafa and Atiya
\protect\hyperlink{ref-abu1996}{1996}).

The ARIMA model is one of the most commonly used methods in financial
forecasting (Fahimifard et al.
\protect\hyperlink{ref-fahimifard2009}{2009}). It is known for its
attractive quality of being a rich class of processes, making it
possible to find a model that adequately fits the data (Fahimifard et
al. \protect\hyperlink{ref-fahimifard2009}{2009}). Some research has
shown that the nature of exchange rate data is non-linear, and this
non-linearity may not be captured adequately by the ARIMA model (Zhang
and Hu \protect\hyperlink{ref-zhang1998}{1998}). Linear parametric
forecasting techniques such as the generalised autoregressive
conditional heteroskedasticity (GARCH) model has been proposed as a
solution to this problem (Zhang and Hu
\protect\hyperlink{ref-zhang1998}{1998}). Although the GARCH model takes
care of the non-linearity in the data, the pre-specification of the
model may prevent it from capturing all the different non-linearities
which may be present in the data.

With the evolution of computational power, other non-linear
non-parametric techniques such as Artificial Neural Networks (ANN),
Support Vector Regression (SVR), Fuzzy Logistic Systems and hybrid
versions of the ARIMA model have been proposed. The ARIMA model has been
used as the standard benchmark model against which these other methods
have been compared. Although they have produced significantly better
results than the ARIMA model, building these models is a complex task.
The ARIMA model is more tractable, less computationally expensive, and
has been used as the building blocks for more advanced models such as
the autoregressive fractionally integrated (ARFIMA), Fuzzy ARIMA, and
seasonal ARIMA (SARIMA) model which have been used for exchange rate
forecasting.

Facebook recently open sourced their forecasting model Prophet. Prophet
is said to be preferable over the ARIMA model as it has the ability to
forecast at scale, is easily interpretable and has parameters that can
be intuitively adjusted (Taylor and Letham
\protect\hyperlink{ref-taylor2017}{2017}). This paper aims to compare
the forecasts produced by the traditional ARIMA model and Prophet
through an evaluation of Bitcoin/ZAR, using a Mincer-Zarnowitz approach
of measuring forecast accuracy. Section 2 will consider the origins of
the ARIMA model and how it has performed and evolved over time by
exploring empirical studies where it has been used for forecasting
exchange rates. Section 3 will also consider other techniques that have
been used to forecast exchange rates and how they have been compared to
the ARIMA model. Section 4 will look at hybrid models that were inspired
by the ARIMA model, while section 5 looks at forecasting with Prophet.
Finally, section 6 will look at the Mincer-Zarnowitz approach to
evaluating forecasts.

\section{Forecasting Exchange Rates with the ARIMA
Model}\label{forecasting-exchange-rates-with-the-arima-model}

\subsection{Birth of ARIMA}\label{birth-of-arima}

ARIMA models became highly popular since its introduction by Box and
Jenkins, when it was shown that they could outperform complex
econometric models in a variety of situations (Hibon and Makridakis
\protect\hyperlink{ref-spyros1997}{1997}). The ARIMA model expresses the
process \{\(y_t\)\} as a function of the weighted average of past values
of the process and lagged values of the residuals. The weighted average
of the past p values of the process represents an autoregressive (AR)
process of order p.~It feeds back past values of the process into the
current value, inducing correlation between all lags of the process. The
weighted average of the q lagged residuals represents a moving average
(MA) process of order q. The purpose of mixing the MA process with the
AR process is to reduce the large number of past values required by AR
processes and to control for the autocorrelation which it creates
between lagged values of the process. The combination of the AR(p) and
MA(q) process results in a more parsimonious model, and forms a
stationary autoregressive moving average (ARMA(p,q)) process defined as:

\[ 
    y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
\] where c is a constant and \{\(e_t\)\} is a white noise process with
zero mean and variance \(\sigma^2\).

The ARIMA(p,d,q) model generalises the ARMA model in that it includes
both stationary and non-stationary processes. The parameter d is the
degree of differencing required to render the process stationary. If d
is equal to zero the process is stationary and equivalent to an ARMA
model, and if d is strictly positive the process requires differencing
to make it stationary. The ARIMA model can be defined succinctly using
the backward shift operator B, which shifts the process back by one unit
of time, and is defined as \(By_t = y_{t-1}\). The ARIMA model has the
form:

\[ 
    (1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
\] where c is a constant and \{\(e_t\)\} is a white noise process with
zero mean and variance \(\sigma^2\).

\subsection{The Box-Jenkins Methodology and Automatic
Selection}\label{the-box-jenkins-methodology-and-automatic-selection}

Box and Jenkins proposed a set of guidelines that can be followed when
selecting ARIMA models. This systematic procedure to designing ARIMA
models has made them highly popular (Hibon and Makridakis
\protect\hyperlink{ref-spyros1997}{1997}). It consists of a four-stage
iterative process in which: 1) the process is either transformed or
differenced to detrend and stabilise the variance of the data, 2) the
autocorrelation and partial autocorrelation plots are used to determine
the order of p and q, 3) the parameters of the model are estimated, 4) a
diagnostic check is performed to ensure that the residuals are a white
noise process. If the residuals are not white noise steps 2-4 are
repeated until a satisfactory model is identified. On the contrary, if
the diagnostic check shows that the residuals are random then the
developed model is the final model used for forecasting.

Due to the large number of forecasts made, it is useful to have an
automatic procedure which is able to select the appropriate ARIMA model
to fit the data. The auto.arima function in R is able to automatically
choose the order of the parameters p, q and d. The order of differencing
is determined by using the KPSS test (Ruppert and Matteson
\protect\hyperlink{ref-ruppert2015}{2015}). The KPSS test checks the
null hypothesis of stationarity and sets d to zero if the null
hypothesis is accepted, otherwise it iteratively increases d by 1 and
tests the null hypothesis until it is accepted (Ruppert and Matteson
\protect\hyperlink{ref-ruppert2015}{2015}). Once the order of
differencing has been determined, the order of p and q are chosen based
on Akaike's Information Criterion (AIC) or the Bayesian Information
Criterion (BIC) (Ruppert and Matteson
\protect\hyperlink{ref-ruppert2015}{2015}).

\subsection{Empirical Applications}\label{empirical-applications}

ARIMA models have been commonly used in financial forecasting and are
popular for observing stock prices and exchange rates due to its power
and statistical properties (C.-S. Lin, Chiu, and Lin
\protect\hyperlink{ref-lin2012}{2012}). They have been used as a
benchmark to compare new forecasting techniques which have emerged over
time. Kamruzzaman and Sarker
(\protect\hyperlink{ref-kamruzzaman2003}{2003}) applied an ARIMA model
to forecast the exchange rate between the Australian dollar and six
other currencies. Its results were used to evaluate the performance of
three Neural Network models. The Normalised Mean Square Error (NMSE) and
Mean Absolute Error (MAE) were used to measure the forecast errors while
the Directional Symmetry (DS), Correct Up trend (CU) and Correct Down
trend (CD) were used to measure the accuracy of the direction of the
forecasts. Similarly, C.-S. Lin, Chiu, and Lin
(\protect\hyperlink{ref-lin2012}{2012}) employed an ARIMA model to
forecast the Taiwan New Dollar exchange rate. The aim of applying the
ARIMA model was to analyse the ability of a proposed hybrid model
created from the Squares Support Vector Regression (LSSVR) and Empirical
Mode Decomposition (EMD) model. C.-S. Lin, Chiu, and Lin
(\protect\hyperlink{ref-lin2012}{2012}) used the same indicators as
Kamruzzaman and Sarker (\protect\hyperlink{ref-kamruzzaman2003}{2003})
to evaluate the accuracy of the direction of the forecasts. To measure
the forecast errors produced they applied the mean absolute percentage
error (MAPE), root‐mean‐square error (RMSE) and mean absolute difference
(MAD). Khashei, Bijari, and Ardali
(\protect\hyperlink{ref-khashei2012}{2012}) also investigated the
ability to forecast exchange rates by considering the British pound and
US dollar, over a short and long forecast horizon. They compared the
predictive capability of their proposed hybrid model consisting of an
ARIMA and a Probabilistic Neural Network (PNN), to the traditional ARIMA
model, and justified the use of the ARIMA as a benchmark, by claiming
that it is the most important linear model. The ARIMA model has also
yielded satisfactory results when predicting exchange rates. Nwankwo
(\protect\hyperlink{ref-nwankwo2014}{2014}) forecasted the Nairo/dollar
rate and diagnostic testing revealed that the ARIMA(1,0,0) model was the
best fit for the data based on the AIC.

The ARIMA model is attractive as it is tractable and produces good
forecasts in short time periods where more than 100 observations are
used (Tseng et al. \protect\hyperlink{ref-tseng2001}{2001}). Although
the ARIMA model has the advantage of ease of implementation and
flexibility, it fails to capture the non-linearity and volatility
present in exchange rate data.

\subsection{Evolution of ARIMA}\label{evolution-of-arima}

Over time, the ARIMA model has evolved to cater for a wider variety of
data that it can be applied to, and to compensate for some of its
shortcomings. The most popular versions of the ARIMA model that has been
implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA)
and Fractional ARIMA model (FARIMA).

The SARIMA model was introduced by Box and Jenkins as a generalisation
of the ARIMA model to capture the periodic behaviour of data. SARIMA
models extend the ARIMA class by including a term for seasonal
differencing. Exchange rate data that has been observed have shown
seasonality, and have been attempted to be fitted by SARIMA models
(Etuk, Wokoma, and Moffat \protect\hyperlink{ref-etuk2013}{2013}). Etuk,
Wokoma, and Moffat (\protect\hyperlink{ref-etuk2013}{2013}) modelled the
Naira/CFA Franc exchange rate which exhibited monthly seasonality using
an additive SARIMA model, to demonstrate that it can be a useful fit to
the data. Their results showed that the SARIMA model adequately
described the variation in the exchange rate series. This agrees with
the results of a study by Etuk and others
(\protect\hyperlink{ref-etuk2012}{2012}) which fitted a multiplicative
SARIMA model to the Naira/Dollar exchange rates.

The ARFIMA model is another generalisation of the ARIMA model in that
the degree required to make the data stationary can assume any real
value, and is no longer restricted to the integer domain. While the
ARIMA model displays exponentially decaying autocorrelation between
observations the further apart they are in time, the ARFIMA model has
the ability to capture the dependence between observations that are
widely spread apart in time (Cheung
\protect\hyperlink{ref-cheung1993}{1993}). This makes the ARFIMA model
more parsimonious since it can capture long memory in data as well as
short term dynamics (Cheung \protect\hyperlink{ref-cheung1993}{1993}).
Cheung (\protect\hyperlink{ref-cheung1993}{1993}) fitted the ARFIMA
model to examine five exchange rates, and found that there was strong
evidence of long memory in the exchange rate time series. Although the
ARFIMA model fitted the data well, it failed to beat the random walk in
out-of sample forecasts.

GARCH models were later developed due to the failure of ARIMA models to
capture the volatility in financial markets (Anastasakis and Mort
\protect\hyperlink{ref-anastasakis2009}{2009}). They have been widely
used to forecast the volatility of stock returns, interest rates and
exchange rates and are useful for portfolio allocation decisions as well
as option pricing (Cermak \protect\hyperlink{ref-cermak2017}{2017}).
Cermak (\protect\hyperlink{ref-cermak2017}{2017}) employed a GARCH(1,1)
model to analyse the volatility of Bitcoin in countries in which Bitcoin
is commonly traded, while Hsieh (1989) applied a GARCH model to
investigate five exchange rates. Hsieh's (1989) results showed that
although the GARCH model outperformed the random walk, some non-linear
information still remained in the residuals. In a similar study,
Fahimifard et al. (\protect\hyperlink{ref-fahimifard2009}{2009})
illustrated that the GARCH model outperformed the ARIMA, while Neural
Networks outperformed both the GARCH and ARIMA model.

\section{Other techniques used to forecast exchange
rates}\label{other-techniques-used-to-forecast-exchange-rates}

Over time, financial forecasting methods have moved away from linear
models like ARIMA and GARCH, to soft computing techniques. These more
complex techniques are non-linear and can fit complex time series more
easily (Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}).
Unlike regression analysis and the ARIMA model, soft computing
techniques do not impose structural assumptions on the model apriori
(Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}). Some of
the most commonly used artificial intelligence methods used to forecast
exchange rate data are Neural Networks and the Fuzzy Logistic Systems.

\subsection{Neural Networks}\label{neural-networks}

Artificial Neural Networks are commonly used in financial forecasting
and are more advantageous than other non-linear forecasting methods.
They can approximate any continuous function, are data driven, and can
adapt to non-stationary environments (Khashei and Bijari
\protect\hyperlink{ref-khashei2011}{2011}). Neural Networks mimic the
structure of the brain and consists of three layers: the input, hidden
and output layer, each of which contain nodes which are representative
of the neurons in our brain. The job of the nodes in the input layer is
to send signals along connections to the nodes in the hidden layer,
where the connections are representative of the synapses in our brain.
The nodes in the hidden layer uses an activation function and assigns a
weight to each connection according to the level of importance of
information. The value in the hidden layer is then sent along
connections to the output layer, where weights are again assigned to
each connection. The error is then computed in the output layer and is
backpropagated through the network. The weights are adjusted according
to the error that they contribute, allowing the ANN to be trained and
learn. This learning process is repeated until the network achieves the
desired output.

ANN's have had many successful applications in forecasting exchange
rates. In a study done by Fahimifard et al.
(\protect\hyperlink{ref-fahimifard2009}{2009}), the ANN was found as an
effective way to improve the forecasts of exchange rates. Superior
results were produced when compared to the ARIMA and GARCH model using
the RMSE, MSE and MAD as measure of performance. Similar results were
found by Zhang and Hu (\protect\hyperlink{ref-zhang1998}{1998}) however
the ANN failed to outperform the random walk over long forecast
horizons. Contrary to the findings above, Kuan and Liu
(\protect\hyperlink{ref-kuan1995}{1995}) attempted to investigate how
well feedforward and recurrent networks capture the non-linearity of
exchange rate data and obtained conflicting results with neural networks
having significantly better market timing ability or lower out-of-sample
Mean Square Prediction Error (MSPE) compared to the ARMA models for some
exchange rates only. Franses and Van Homelen
(\protect\hyperlink{ref-franses1998}{1998}) suggested that the
non-linear feature of exchange rates that are picked up by ANNs may
actually be due to neglected GARCH effects. They investigated the
application of ANNs to forecast exchange rates and found that the
presence of GARCH in the data may mistakenly lead one to believe that
the returns can be forecasted on the exchange rates themselves. Their
results show that there is no gain in producing out-of-sample forecasts
using ANN if the data is not truly non-linear.

Although neural networks have been broadly applied in financial
forecasting, the process of building neural networks is a complex task
and there is no consistent method to design them compared to the
traditional Box-Jenkins ARIMA model. Unlike ARIMA models, the
performance of ANNs is sensitive to many modelling factors such as the
number of input nodes and the training sample size (Zhang and Hu
\protect\hyperlink{ref-zhang1998}{1998}).

\subsection{Fuzzy Logistic Systems}\label{fuzzy-logistic-systems}

Fuzzy Logistic Systems were initially developed to solve problems
involving linguistic terms, and have been used successful in financial
forecasting (Khashei, Bijari, and Ardali
\protect\hyperlink{ref-khashei2009}{2009}). Unlike computers, the human
decision-making process includes a range of possibilities spanning `yes'
and `no'. Fuzzy logic tries to imitate this reasoning and
decision-making process by taking a discrete decision space that
consists of a large number of states and turning it into a continuous
space. This allows for more finer decisions to be provided as opposed to
discrete decisions.

Santos, Costa, and Santos Coelho
(\protect\hyperlink{ref-santos2007}{2007}) investigated how well fuzzy
systems and neural networks perform compared to the traditional ARMA and
GARCH model. He examined the forecasts of Brazilian exchange rate
returns looking at different frequencies of the series and compared
their one-step-ahead forecasts. By analysing the RMSE, U-Theil
inequality index, percentage of corrected predicted signals (CPS), and
the Pesaran-Timmermann (PT) predictive failure statistic, he found that
the Fuzzy Systems and Neural Networks achieved higher returns based on
the forecasts they produced. Similar results were found by Khashei,
Bijari, and Ardali (\protect\hyperlink{ref-khashei2009}{2009}) when he
analysed the predictive capabilities of Fuzzy Systems, Neural Networks,
the traditional ARIMA model, and a Fuzzy ARIMA model. The MAE and MSE of
the ARIMA model were higher compared to the other models.
